### 2024-01-22 parts/correction.tex
update proofs and paper

\begin{lemma}
    \label{n-poly-combinatorics:lem}
    Let $f$ be a  \kl{$\Nat$-polyregular} function. 
    There exists $\omega \in \Nat$
    such that for all $p \in \Nat$,
    for all $\alpha_0, \dots, \alpha_p \in \Sigma^*$,
    for all $u_1, \dots, u_p \in \Sigma^*$,
    there exists a polynomial $P \in \Rel[X_1, \dots, X_p]$
    whose \kl{maximal monomials} are \kl{non-negative},
    and such that for all $X_1, \dots, X_p \geq \omega$:
    \begin{equation*}
        f\left(
            \alpha_0 \prod_{i = 1}^p u_i^{\omega \times X_i} \alpha_i
        \right)
        = P(X_1, \dots, X_p) \quad .
    \end{equation*}
\end{lemma}
\begin{proof}
    Let $\omega$ be an idempotent power for the finite monoid $M$,
    and
    $w \defined \alpha_0 \prod_{i = 1}^p (u_i^{\omega \times X_i} \alpha_i)$.
    Let $k \in \Nat$, and $\pi \colon M^k \topartial \Nat$ be such that
    $f = \pi^\dagger$. That is, 
    $f(w)$ is the sum over all factorizations of $w$
    into $k$ words $(w_1, \dots, w_k) \in \Sigma^*$
    of the value $\pi(\mu(w_1), \dots, \mu(w_k))$.

@@

@@

\begin{lemma}
    \label{derivation-simplifies:lemma}
    For all $P \in \Rel[\vec{X}]$ that are non-constant,
    for all $K \in \Nat$,
    $\MaximalMonomials(\Diff{K}{P}) \hoarele
    \MaximalMonomials(P)$.
\end{lemma}
\begin{proof}
    By linearity of $\Diff{K}{P}$ (\cref{discrete-deriv-linear:fact}), 
    it suffices to prove the result
    for \kl{monomials} $S \in \Monomials$.
\end{proof}

\begin{fact}
    Let $K > 0$,
    and let $M,T$ be two \kl{monomials}, such that
    $T$ \kl{strictly divides} $M$.
    Then,
    $T$ \kl{divides} some \kl{maximal monomial}
    of $\Diff{K}{M}$.
\end{fact}

\begin{lemma}
    \label{derivation-covers:lemma}
    For all $P,Q \in \Rel[\vec{X}]$ that are non-constant,
    for all $K > 0$,
    if $\MaximalMonomials(P) \hoarele \MaximalMonomials(Q)$,
    then
    $\MaximalMonomials(P) \hoareleq \MaximalMonomials(\Diff{K}{Q})$.
\end{lemma}

\begin{lemma}
    \label{derivation-stabilises-correct:lem}
    Let $P \in \CorrectPoly$,
    $P_1$ be the sum of \kl{maximal monomials} of $P$,
    and $P_2 \defined P - P_1$ be the sum of
    non-maximal monomials of $P$.
    There exists a computable $K$,
    such that
    $(\Diff{K}{P_1} + P_2) \in \CorrectPoly$.
\end{lemma}

>>>>

\begin{lemma}
    \label{n-poly-combinatorics:lem}
    Let $f$ be a  \kl{$\Nat$-polyregular} function. 
    There exists $\omega \in \Nat$
    such that for all $p \in \Nat$,
    for all $\alpha_0, \dots, \alpha_p \in \Sigma^*$,
    for all $u_1, \dots, u_p \in \Sigma^*$,
    there exists a polynomial $P \in \Rel[X_1, \dots, X_p]$
    whose \kl{maximal monomials} are \kl{non-negative},
    and such that for all $X_1, \dots, X_p \geq \omega$:
    \begin{equation*}
        f\left(
            \alpha_0 \prod_{i = 1}^p u_i^{\omega X_i} \alpha_i
        \right)
        = P(X_1, \dots, X_p) \quad .
    \end{equation*}
\end{lemma}
\begin{proof}
    Let $\omega$ be an idempotent power for the finite monoid $M$,
    and
    $w \defined \alpha_0 \prod_{i = 1}^p (u_i^{\omega \times X_i} \alpha_i)$.
    Let $k \in \Nat$, and $\pi \colon M^k \topartial \Nat$ be such that
    $f = \pi^\dagger$. That is, 
    $f(w)$ is the sum over all factorizations of $w$
    into $k$ words $(w_1, \dots, w_k) \in \Sigma^*$
    of the value $\pi(\mu(w_1), \dots, \mu(w_k))$.

@@

\begin{fact}
    \label{maxi-monomials-submodular:fact}
    Let $P, Q$ be two polynomials.
    Then, $\MaximalMonomials(P+Q) \hoareleq \MaximalMonomials(P)
    \cup \MaximalMonomials(Q)$.
\end{fact}

@@

\begin{fact}
    \label{derivation-monomials:fact}
    Let $K > 0$,
    and let $M,T$ be two \kl{monomials}, such that
    $T$ \kl{strictly divides} $M$.
    Then,
    all \kl{monomials} of $\Diff{K}{M}$ \kl{strictly divides} $M$,
    and 
    $T$ \kl{divides} some \kl{maximal monomial}
    of $\Diff{K}{M}$, which has a coefficient
    that is a multiple of $K$.
\end{fact}
\begin{proof}
    Let $M \defined X_1^{k_1} \cdots X_n^{k_n}$,
    where $k_1, \dots, k_n > 0$ and $n > 0$.
    Let us compute
    \begin{align*}
        \Diff{K}{M} &= \prod_{i = 1}^n X_i^{k_i} - \prod_{i = 1}^n (X_i - K)^{k_i} \\
                    &= \sum_{p = 1}^n \frac{K}{X_p - K} \prod_{i = 1}^n (X_i - K)^{k_i} 
    \end{align*}
    Therefore, 
    the \kl{maximal monomials} of $\Diff{K}{M}$
    are precisely
    of the form $K \frac{\prod_{i = 1}^n X_i^{k_i}}{X_p}$ for some $1 \leq p \leq i$,
    which \kl{strictly divide} $M$.
    Let $T$ be a \kl{monomial} that \kl{strictly divides} $M$,
    then, $T = \prod_{i = 1}^n X_i^{t_i}$, 
    for all $1 \leq i \leq n$, $t_i \leq k_i$,
    and there exists $1 \leq p \leq n$ such that$ t_j < k_j$.
    In particular, $T$ \kl{divides}
    $K \frac{\prod_{i = 1}^n X_i^{k_i}}{X_p}$.
\end{proof}

\begin{lemma}
    \label{derivation-simplifies:lemma}
    For all $P \in \Rel[\vec{X}]$ that are non-constant,
    for all $K \in \Nat$,
    $\MaximalMonomials(\Diff{K}{P}) \hoarele
    \MaximalMonomials(P)$.
\end{lemma}
\begin{proof}
    The result holds for \kl{monomials}
    thanks to \cref{derivation-monomials:fact}.
    It extends by linearity to polynomials 
    thanks to \cref{maxi-monomials-submodular:fact,discrete-deriv-linear:fact}.
\end{proof}

\begin{lemma}
    \label{derivation-stabilises-correct:lem}
    Let $P \in \CorrectPoly$,
    $P_1$ be the sum of \kl{maximal monomials} of $P$,
    and $P_2 \defined P - P_1$ be the sum of
    non-maximal monomials of $P$.
    There exists a computable $K$,
    such that
    $Q \defined (\Diff{K}{P_1} + P_2) \in \CorrectPoly$.
\end{lemma}
\begin{proof}
    If $P$ is constant, then $P_2 = 0$, and $\Diff{K}{P_1} = 0$,
    hence for $K = 0$, $Q \in \CorrectPoly$.
    Otherwise, let $\nu \colon \vec{X} \topartial \Nat$,
    and let us analyze the \kl{maximal monomials} $T$
    of $\restr{Q}{\nu}$.
    Because $T$ is a \kl{monomial} of $\restr{Q}{\nu}$,
    it \kl{divides} some \kl{monomial} in $Q$, and
    using \cref{derivation-monomials:fact},
    it \kl{divides} some \kl{monomial} $M$ in 
    $\Diff{K}{P_1}$.
    Now, the monomial $T$ is computed as follows:
    \begin{align*}
        T &= 
        \sum_{S \in \Monomials(Q), T \divides S}
        \restr{S}{\nu} \\
          &= 
          \sum_{S \in \Monomials(\Diff{K}{P_1}), T \divides S}
            \restr{S}{\nu}
        + 
          \sum_{S \in \Monomials(P_2), T \divides S}
            \restr{S}{\nu}
        \\
          &\geq
          \sum_{S \in \Monomials(\Diff{K}{P_1}), T \divides S}
            \restr{S}{\nu}
          - 
            \left|
          \sum_{S \in \Monomials(P_2), T \divides S}
            \restr{S}{\nu}
            \right|
    \end{align*}

\end{proof}
